

% Chapter 1

\chapter{The way of the program  ｜  程序之道}

The goal of this book is to teach you to think like a computer
scientist.  This way of thinking combines some of the best features of
mathematics, engineering, and natural science.  Like mathematicians,
computer scientists use formal languages to denote ideas (specifically
computations).  Like engineers, they design things, assembling
components into systems and evaluating tradeoffs among alternatives.
Like scientists, they observe the behavior of complex systems, form
hypotheses, and test predictions.

本书的目标是教你像计算机科学家一样思考。这一思考方式集成了数学、工程以及自然科学的一些最好的特点。像数学家一样，计算机科学家使用形式语言表示思想（具体来说是计算）。像工程师一样，计算机科学家设计东西，将零件组成系统，在各种选择之间寻求平衡。像科学家一样，计算机科学家观察复杂系统的行为，形成假设并且对预测进行检验。
\index{problem solving}  \index{问题求解}

The single most important skill for a computer scientist is {\bf
  problem solving}.  Problem solving means the ability to formulate
problems, think creatively about solutions, and express a solution
clearly and accurately.  As it turns out, the process of learning to
program is an excellent opportunity to practice problem-solving
skills.  That's why this chapter is called, ``The way of the
program''.

对于计算机科学家，最重要的技能是\textbf{问题求解}的能力 。问题求解（problem solving）意味着对问题进行形式化，寻求创新型的解决方案，并且清晰、准确地表达解决方案的能力。事实证明，学习编程的过程是锻炼问题解决能力的一个绝佳机会。这就是为什么本章被称为“程序之道”。

On one level, you will be learning to program, a useful skill by
itself.  On another level, you will use programming as a means to an
end.  As we go along, that end will become clearer.

一方面，你将学习如何编程，这本身就是一个有用的技能。另一方面，你将把编程作为实现自己目的的手段。随着学习的深入，你会更清楚自己的目的。

\section{What is a program?  |  什么是程序？}

A {\bf program} is a sequence of instructions that specifies how to
perform a computation.  The computation might be something
mathematical, such as solving a system of equations or finding the
roots of a polynomial, but it can also be a symbolic computation, such
as searching and replacing text in a document or something
graphical, like processing an image or playing a video.

\emph{程序}是一系列说明如何执行计算（computation）的指令。计算可以是数学上的计算，例如寻找公式的解或多项式的根，也可以是一个符号计算（symbolic computation），例如在文档中搜索并替换文本或者图片，就像处理图片或播放视频。
\index{program}

The details look different in different languages, but a few basic
instructions appear in just about every language:

不同编程语言中，程序的具体细节也不一样，但是有一些基本的指令几乎出现在每种语言当中：

\begin{description}

\item[input:] Get data from the keyboard, a file, the network, or some
other device.

\item[output:] Display data on the screen, save it in a
file, send it over the network, etc.

\item[math:] Perform basic mathematical operations like addition and
multiplication.

\item[conditional execution:] Check for certain conditions and
run the appropriate code.

\item[repetition:] Perform some action repeatedly, usually with
some variation.

\end{description}

\begin{description}

\item[输入（input）：] 从键盘、文件、网络或者其他设备获取数据。

\item[输出（output）：] 在屏幕上显示数据，将数据保存至文件，通过网络传送数据，等等。

\item[数学（math）：] 执行基本的数学运算，如加法和乘法。

\item[有条件执行（conditional execution）:] 检查符合某个条件后，执行相应的代码。

\item[重复（repetition）:] 检查符合某个条件后，执行相应的代码。

\end{description}

Believe it or not, that's pretty much all there is to it.  Every
program you've ever used, no matter how complicated, is made up of
instructions that look pretty much like these.  So you can think of
programming as the process of breaking a large, complex task
into smaller and smaller subtasks until the subtasks are
simple enough to be performed with one of these basic instructions.

无论你是否相信，这几乎是程序的全部指令了。每个你曾经用过的程序，无论多么复杂，都是由跟这些差不多的指令构成的。因此，你可以认为编程就是将庞大、复杂的任务分解为越来越小的子任务，直到这些子任务简单到可以用这其中的一个基本指令执行。

\section{Running Python  |  运行Python}

One of the challenges of getting started with Python is that you
might have to install Python and related software on your computer.
If you are familiar with your operating system, and especially
if you are comfortable with the command-line interface, you will
have no trouble installing Python.  But for beginners, it can be
painful to learn about system administration and programming at the
same time.

Python入门的一个障碍，是你可能需要在电脑上安装Python和相关软件。
如果你能熟练使用命令行 (command-line interface) ，安装Python对你来说就不是问题了。但是对于初学者，同时学习系统管理 (system administration) 和编程这两方面的知识是件痛苦的事。
\index{running Python}  \index{Python!running}

To avoid that problem, I recommend that you start out running Python
in a browser.  Later, when you are comfortable with Python, I'll
make suggestions for installing Python on your computer.

为了避免这个问题，我建议你首先在浏览器中运行Python。等你对Python更加了解之后，我会建议你在电脑上安装Python。
\index{Python in a browser}

There are a number of web pages you can use to run Python.  If you
already have a favorite, go ahead and use it.  Otherwise I recommend
PythonAnywhere.  I provide detailed instructions for getting started
at \url{http://tinyurl.com/thinkpython2e}.

网络上有许多网页可以让你运行Python。如果你已经有最喜欢的网站，那就打开网页运行Python吧。如果没有，我推荐PythonAnywhere。我在 \href{http://tinyurl.com/thinkpython2e}{http://tinyurl.com/thinkpython2e} 给出了详细的使用指南。
\index{PythonAnywhere}

There are two versions of Python, called Python 2 and Python 3.
They are very similar, so if you learn one, it is easy to switch
to the other.  In fact, there are only a few differences you will
encounter as a beginner.
This book is written for Python 3, but I include some notes
about Python 2.

目前Python有两个版本，分别是Python 2和Python 3。二者十分相似，因此如果你学过某个版本，可以很容易地切换到另一个版本。事实上，作为初学者，你只会接触到很少数的不同之处。
本书采用的是Python 3，但是我会加入一些关于Python 2的说明。
\index{Python 2}

The Python {\bf interpreter} is a program that reads and executes
Python code.  Depending on your environment, you might start the
interpreter by clicking on an icon, or by typing {\tt python} on
a command line.
When it starts, you should see output like this:

Python的 \textbf{解释器} 是一个读取并执行 Python 代码的程序。  根据你的电脑环境不同，你可以通过双击图标，或者在命令行输入\lstinline{python} 的方式来启动解释器。解释器启动后，你应该看到类似下面的输出：
\index{interpreter}

\begin{lstlisting}
Python 3.4.0 (default, Jun 19 2015, 14:20:21)
[GCC 4.8.2] on linux
Type "help", "copyright", "credits" or "license" for more information.
>>>
\end{lstlisting}

%
The first three lines contain information about the interpreter
and the operating system it's running on, so it might be different for
you.  But you should check that the version number, which is
{\tt 3.4.0} in this example, begins with 3, which indicates that
you are running Python 3.  If it begins with 2, you are running
(you guessed it) Python 2.

前三行中包含了关于解释器及其运行的操作系统的信息，因此你看到的内容可能不一样。但是你应该检查下版本号是否以3开头，上面示例中的版本号是3.4.0。如果以3开头，那说明你正在运行Python 3。如果以2开头，那说明你正在运行（你猜对了）Python 2。

The last line is a {\bf prompt} that indicates that the interpreter is
ready for you to enter code.
If you type a line of code and hit Enter, the interpreter displays the
result:

最后一行是一个提示符（prompt），表明你可以在解释器中输入代码了。如果你输入一行代码然后按回车（Enter），解释器就会显示结果：
\index{prompt}

\begin{lstlisting}
>>> 1 + 1
2
\end{lstlisting}
%
Now you're ready to get started.
From here on, I assume that you know how to start the Python
interpreter and run code.

现在你已经做好了开始学习的准备。接下来，我将默认你已经知道如何启动Python解释器和执行代码。


\section{The first program  |  第一个程序}
\label{hello}
\index{Hello, World}

Traditionally, the first program you write in a new language
is called ``Hello, World!'' because all it does is display the
words ``Hello, World!''.  In Python, it looks like this:

根据传统，你用一门新语言写的第一个程序叫做 ``Hello, World!'' ，因为它的功能只不过是显示单词 ``Hello, World!'' 。在Python中，它看起来是这样：

\begin{lstlisting}
>>> print('Hello, World!')
\end{lstlisting}
%
This is an example of a {\bf print statement}, although it
doesn't actually print anything on paper.  It displays a result on the
screen.  In this case, the result is the words:


这是一个 ``print'' 函数的示例，尽管它并不会真的在纸上打印。  它将结果显示在屏幕上。在此例中，结果是单词：

\begin{lstlisting}
Hello, World!
\end{lstlisting}
%
The quotation marks in the program mark the beginning and end
of the text to be displayed; they don't appear in the result.

程序中的单引号标记了被打印文本的首尾；它们不会出现在结果中。
\index{quotation mark}  \index{print statement}  \index{statement!print}

The parentheses indicate that {\tt print} is a function.  We'll get
to functions in Chapter~\ref{funcchap}.
\index{function} \index{print function}

In Python 2, the print statement is slightly different; it is not
a function, so it doesn't use parentheses.
\index{Python 2}

括号说明 print 是一个函数。我们将在第三章介绍函数。在Python 2中， print是一个语句；不是函数，所以不需要使用括号。

\begin{lstlisting}
>>> print 'Hello, World!'
\end{lstlisting}
%
This distinction will make more sense soon, but that's enough to
get started.

很快你就会明白二者之间的区别\footnote{译注：Python核心开发者Brett Cannon详细解释了\href{http://codingpy.com/article/why-print-became-a-function-in-python-3/}{为什么print在Python 3中变成了函数}。}，现在知道这些就足够了。


\section{Arithmetic operators  |  算术运算符}
\index{operator!arithmetic}  \index{arithmetic operator}

After ``Hello, World'', the next step is arithmetic.  Python provides
{\bf operators}, which are special symbols that represent computations
like addition and multiplication.

The operators \lstinline{+}, \lstinline{-}, and \lstinline{*} perform addition,
subtraction, and multiplication, as in the following examples:

接下来介绍算术。Python提供了许多代表加法和乘法等运算的特殊符号，叫做 \emph{运算符} (operators)。

运算符 \lstinline{+} 、\lstinline{-} 和 \lstinline{*} 分别执行加法、减法和乘法，详见以下示例：

\begin{lstlisting}
>>> 40 + 2
42
>>> 43 - 1
42
>>> 6 * 7
42
\end{lstlisting}
%
The operator \lstinline{/} performs division:

运算符 \lstinline{/} 执行除法运算：


\begin{lstlisting}
>>> 84 / 2
42.0
\end{lstlisting}
%
You might wonder why the result is {\tt 42.0} instead of {\tt 42}.
I'll explain in the next section.

Finally, the operator {\tt **} performs exponentiation; that is,
it raises a number to a power:

你可能会问，为什么结果是 \lstinline{42.0} ，而不是 \lstinline{42} 。在下节中，我会进行解释。

% todo: add link here

最后，运算符 \lstinline{*} 执行乘方运算；也就是说，它将某个数字乘以自身相应的次数：

\begin{lstlisting}
>>> 6**2 + 6
42
\end{lstlisting}
%
In some other languages, \verb"^" is used for exponentiation, but
in Python it is a bitwise operator called XOR.  If you are not
familiar with bitwise operators, the result will surprise you:

某些语言使用 \lstinline{^} 运算符执行乘方运算，但是在Python中，它却属于一种位运算符，叫做 XOR 。  如果你对位运算符不太了解，那么下面的结果会让你感到惊讶：

\begin{lstlisting}
>>> 6 ^ 2
4
\end{lstlisting}
%
I won't cover
bitwise operators in this book, but you can read about
them at \url{http://wiki.python.org/moin/BitwiseOperators}.

我们不会在本书中过深涉及位运算符，你可以通过阅读 \href{http://wiki.python.org/moin/BitwiseOperators}{Python百科 --- 位运算符} ，了解相关内容。
\index{bitwise operator}  \index{operator!bitwise}
\index{位运算符}  \index{运算符!位运算符}

%%
\section{Values and types  |  值和类型}
\index{value}  \index{type}  \index{string}


A {\bf value} is one of the basic things a program works with, like a
letter or a number.  Some values we have seen so far are {\tt 2},
{\tt 42.0}, and \verb"'Hello, World!'".

These values belong to different {\bf types}:
{\tt 2} is an {\bf integer}, {\tt 42.0} is a {\bf floating-point number},
and \verb"'Hello, World!'" is a {\bf string},
so-called because the letters it contains are strung together.
\index{integer}
\index{floating-point}

\emph{值} (value) 是程序处理的基本数据之一，一个单词或一个数字都是值的实例。 我们目前已经接触到的值有：\lstinline{2}，\lstinline{42.0}，和 \lstinline{'Hello World!'}。

这些值又属于不同的 \emph{类型} (types) ：\lstinline{2} 是一个 \emph{整型数} (integer)，\lstinline{42.0} 是一个 \emph{浮点数} (floating point number)，而 \lstinline{'Hello, World!'}  则是一个  \emph{字符串} (string)，这么称呼是因为其中的字符被串在了一起 (strung together)。

If you are not sure what type a value has, the interpreter can
tell you:

如果你不确定某个值的类型是什么，\emph{解释器}可以告诉你：

\begin{lstlisting}
>>> type(2)
<class 'int'>
>>> type(42.0)
<class 'float'>
>>> type('Hello, World!')
<class 'str'>
\end{lstlisting}
%
In these results, the word ``class'' is used in the sense of
a category; a type is a category of values.
\index{class}

Not surprisingly, integers belong to the type {\tt int},
strings belong to {\tt str} and floating-point
numbers belong to {\tt float}.
\index{type}  \index{string type}  \index{type!str}
\index{int type}  \index{type!int}  \index{float type}
\index{type!float}

What about values like \verb"'2'" and \verb"'42.0'"?
They look like numbers, but they are in quotation marks like
strings.
\index{quotation mark}

``class'' 一词在上面的输出结果中，是类别的意思；一个类型就是一个类别的值。

不出意料，整型数属于 \lstinline{int} 类型，字符串属于 \lstinline{str} 类型，浮点数属于 \lstinline{float} 类型。

那么像 \lstinline{'2'} 和 \lstinline{'42.0'} 这样的值呢？它们看上去像数字，但是又和字符串一样被引号括在了一起？

\begin{lstlisting}
>>> type('2')
<class 'str'>
>>> type('42.0')
<class 'str'>
\end{lstlisting}
%
They're strings.

它们其实是字符串。

When you type a large integer, you might be tempted to use commas
between groups of digits, as in {\tt 1,000,000}.  This is not a
legal {\em integer} in Python, but it is legal:

当你输入一个大数值的整型数时，你可能会想用逗号进行区分，比如说像这样：\lstinline{1,000,000}。在Python中，这不是一个合法的\textbf{整型数}，但是确实合法的值。

\begin{lstlisting}
>>> 1,000,000
(1, 0, 0)
\end{lstlisting}
%
That's not what we expected at all!  Python interprets {\tt
  1,000,000} as a comma-separated sequence of integers.  We'll learn
more about this kind of sequence later.
\index{sequence}

结果和我们预料的完全不同！ Python 把 \lstinline{1,000,000} 当作成了一个以逗号区分的整型数序列。在后面的章节中，我们会介绍更多有关这种序列的知识。

%This is the first example we have seen of a semantic error: the code
%runs without producing an error message, but it doesn't do the
%``right'' thing.
%\index{semantic error}
%\index{error!semantic}
%\index{error message}
% TODO: use this as an example of a semantic error later



\section{Formal and natural languages  |  形式语言和自然语言}
\index{formal language}  \index{natural language}  \index{language!formal}
\index{language!natural}

{\bf Natural languages} are the languages people speak,
such as English, Spanish, and French.  They were not designed
by people (although people try to impose some order on them);
they evolved naturally.

\emph{自然语言} (natural language) 是人们交流所使用的语言，例如英语、西班牙语和法语。它们不是人为设计出来的（尽管有人试图这样做）；而是自然演变而来。

{\bf Formal languages} are languages that are designed by people for
specific applications.  For example, the notation that mathematicians
use is a formal language that is particularly good at denoting
relationships among numbers and symbols.  Chemists use a formal
language to represent the chemical structure of molecules.  And
most importantly:

\emph{形式语言} (formal languages)是人类为了特殊用途而设计出来的。例如，数学家使用的记号 (notation) 就是形式语言，特别擅长表示数字和符号之间的关系。化学家使用形式语言表示分子的化学结构。 最重要的是：

\begin{quote}
{\bf Programming languages are formal languages that have been
designed to express computations.}
\end{quote}

\begin{quote}
\textbf{编程语言是被设计用于表达计算的形式语言。}
\end{quote}

Formal languages tend to have strict {\bf syntax} rules that
govern the structure of statements.
For example, in mathematics the statement
$3 + 3 = 6$ has correct syntax, but
$3 + = 3 \$ 6$ does not.  In chemistry
$H_2O$ is a syntactically correct formula, but $_2Zz$ is not.

形式语言通常拥有严格的 \emph{语法} 规则，规定了详细的语句结构。例如， $3 + 3 = 6$ 是语法正确的数学表达式，而 $3 + = 3 \$ 6$ 则不是； $H_2O$ 是语法正确的化学式，而 $_2Zz$ 则不是。
\index{syntax}

Syntax rules come in two flavors, pertaining to {\bf tokens} and
structure.  Tokens are the basic elements of the language, such as
words, numbers, and chemical elements.  One of the problems with
$3 += 3 \$ 6$ is that \( \$ \) is not a legal token in mathematics
(at least as far as I know).  Similarly, $_2Zz$ is not legal because
there is no element with the abbreviation $Zz$.

语法规则有两种类型，分别涉及\emph{记号} (tokens)和结构。  记号是语言的基本元素，例如单词、数字和化学元素。  $3 + = 3 \$ 6$ 这个式子的问题之一，就是 \$ 在数学中不是一个合法的记号 （至少据我所知）。类似的， $_2Zz$ 也不合法，因为没有一个元素的简写是  $Zz$ 。
\index{token}  \index{structure}

The second type of syntax rule pertains to the way tokens are
combined.  The equation $3 += 3$ is illegal because even though $+$
and $=$ are legal tokens, you can't have one right after the other.
Similarly, in a chemical formula the subscript comes after the element
name, not before.

第二种语法规则与标记的组合方式有关。 $3 + = 3$ 这个方程是非法的，因为即使 $+$ 和 $=$ 都是合法的记号，但是你却不能把它们俩紧挨在一起。类似的，在化学式中，下标位于元素之后，而不是之前。

This is @ well-structured Engli\$h
sentence with invalid t*kens in it.  This sentence all valid tokens
has, but invalid structure with. \footnote{译注：上面两句英文都是不符合语法的，一个包含非法标记，另一个结构不符合语法。}

When you read a sentence in English or a statement in a formal
language, you have to figure out the structure
(although in a natural language you do this subconsciously).  This
process is called {\bf parsing}.

当你读一个用英语写的句子或者用形式语言写的语句时，你都必须要理清各自的结构（尽管在阅读自然语言时，你是下意识地进行的）。这个过程被称为 \emph{解析} (parsing)。
\index{parse}  \index{解析}

Although formal and natural languages have many features in
common---tokens, structure, and syntax---there are some
differences:

虽然形式语言和自然语言有很多共同点——标记、结构和语法，它们也有一些不同：
\index{ambiguity}  \index{redundancy}  \index{literalness}

\begin{description}

\item[ambiguity:] Natural languages are full of ambiguity, which
people deal with by using contextual clues and other information.
Formal languages are designed to be nearly or completely unambiguous,
which means that any statement has exactly one meaning,
regardless of context.

\item[redundancy:] In order to make up for ambiguity and reduce
misunderstandings, natural languages employ lots of
redundancy.  As a result, they are often verbose.  Formal languages
are less redundant and more concise.

\item[literalness:] Natural languages are full of idiom and metaphor.
If I say, ``The penny dropped'', there is probably no penny and
nothing dropping (this idiom means that someone understood something
after a period of confusion).  Formal languages
mean exactly what they say.

\end{description}

\begin{description}

\item[歧义性 (ambiguity)：] 自然语言充满歧义，人们使用上下文线索以及其它信息处理这些歧义。形式语言被设计成几乎或者完全没有歧义，这意味着不管上下文是什么，任何语句都只有一个意义。

\item[冗余性 (redundancy)：] 为了弥补歧义性并减少误解，自然语言使用很多冗余。结果，自然语言经常很冗长。形式语言则冗余较少，更简洁。

\item[字面性 (literalness)：] 自然语言充满成语和隐喻。如果我说``The penny dropped''，可能根本没有便士、也没什么东西掉下来（这个成语的意思是，经过一段时间的困惑后终于理解某事）。形式语言的含义，与它们字面的意思完全一致。

\end{description}

Because we all grow up speaking natural languages, it is sometimes
hard to adjust to formal languages.  The difference between formal and
natural language is like the difference between poetry and prose, but
more so:

由于我们都是说着自然语言长大的，我们有时候很难适应形式语言。形式语言与自然语言之间的不同，类似诗歌与散文之间的差异，而且更加明显：
\index{poetry} \index{prose}

\begin{description}

\item[Poetry:] Words are used for their sounds as well as for
their meaning, and the whole poem together creates an effect or
emotional response.  Ambiguity is not only common but often
deliberate.

\item[Prose:] The literal meaning of words is more important,
and the structure contributes more meaning.  Prose is more amenable to
analysis than poetry but still often ambiguous.

\item[Programs:] The meaning of a computer program is unambiguous
and literal, and can be understood entirely by analysis of the
tokens and structure.

\end{description}

\begin{description}

\item[诗歌 (Poetry)：] 单词的含义和声音都有作用， 整首诗作为一个整理，会对人产生影响，或是引发情感上的共鸣。 歧义不但常见，而且经常是故意为之。

\item[散文 (Prose)：] 单词表面的含义更重要，句子结构背后的寓意更深。 散文比诗歌更适合分析，但仍然经常有歧义。

\item[程序 (Programs)：] 计算机程序的含义是无歧义、无引申义的， 通过分析程序的标记和结构，即可完全理解。

\end{description}

Formal languages are more dense
than natural languages, so it takes longer to read them.  Also, the
structure is important, so it is not always best to read
from top to bottom, left to right.  Instead, learn to parse the
program in your head, identifying the tokens and interpreting the
structure.  Finally, the details matter.  Small errors in
spelling and punctuation, which you can get away
with in natural languages, can make a big difference in a formal
language.

形式语言要比自然语言更加稠密，因此阅读起来花的时间会更长。另外，形式语言的结构也很重要，所以从上往下、从左往右阅读，并不总是最好的策略。相反，你得学会在脑海里分析一个程序，识别不同的标记并理解其结构。最后，注重细节。拼写和标点方面的小错误在自然语言中无伤大雅，但是在形式语言中却会产生很大的影响。

\section{Debugging  |  调试}
\index{debugging}

Programmers make mistakes.  For whimsical reasons, programming errors
are called {\bf bugs} and the process of tracking them down is called
{\bf debugging}.

程序员都会犯错。由于比较奇怪的原因，编程错误被称为 \emph{故障} \footnote{译注：英文为 bug，一般指虫子}，追踪错误的过程被称为 \emph{调试} (debugging) 。

\index{debugging}  \index{bug}

Programming, and especially debugging, sometimes brings out strong
emotions.  If you are struggling with a difficult bug, you might
feel angry, despondent, or embarrassed.

编程，尤其是调试，有时会让人动情绪。如果你有个很难的bug解决不了，你可能会感到愤怒、沮丧抑或是难堪。

There is evidence that people naturally respond to computers as if
they were people.  When they work well, we think
of them as teammates, and when they are obstinate or rude, we
respond to them the same way we respond to rude,
obstinate people (Reeves and Nass, {\it The Media
    Equation: How People Treat Computers, Television, and New Media
    Like Real People and Places}).

有证据表明，人们很自然地把计算机当人来对待。当计算机表现好的时候，我们认为它们是队友，而当它们固执或无礼的时候，我们也会像对待固执或无礼人的一样对待它们(Reeves and Nass, {\it The Media Equation: How People Treat Computers, Television, and New Media Like Real People and Places})。
\index{debugging!emotional response}  \index{emotional debugging}

Preparing for these reactions might help you deal with them.
One approach is to think of the computer as an employee with
certain strengths, like speed and precision, and
particular weaknesses, like lack of empathy and inability
to grasp the big picture.

对这些反应做好准备有助于你对付它们。 一种方法是将计算机看做是一个雇员，拥有特定的长处， 例如速度和精度，也有些特别的缺点，像缺乏沟通以及不善于把握大局。

Your job is to be a good manager: find ways to take advantage
of the strengths and mitigate the weaknesses.  And find ways
to use your emotions to engage with the problem,
without letting your reactions interfere with your ability
to work effectively.

你的工作是当一个好的管理者：找到充分利用优点、摒弃弱点的方法。 并且找到使用你的情感来解决问题的方法， 而不是让你的情绪干扰你有效工作的能力。

Learning to debug can be frustrating, but it is a valuable skill
that is useful for many activities beyond programming.  At the
end of each chapter there is a section, like this one,
with my suggestions for debugging.  I hope they help!

学习调试可能很令人泄气， 但是它对于许多编程之外的活动也是一个非常有价值的技能。 在每一章的结尾，我都会花一节内容介绍一些调试建议，比如说这一节。希望能帮到你！

\section{Glossary  |  术语表}

\begin{description}

\item[problem solving:]  The process of formulating a problem, finding
a solution, and expressing it.
\index{problem solving}

\item[high-level language:]  A programming language like Python that
is designed to be easy for humans to read and write.
\index{high-level language}

\item[low-level language:]  A programming language that is designed
to be easy for a computer to run; also called ``machine language'' or
``assembly language''.
\index{low-level language}

\item[portability:]  A property of a program that can run on more
than one kind of computer.
\index{portability}

\item[interpreter:]  A program that reads another program and executes
it
\index{interpret}

\item[prompt:] Characters displayed by the interpreter to indicate
that it is ready to take input from the user.
\index{prompt}

\item[program:] A set of instructions that specifies a computation.
\index{program}

\item[print statement:]  An instruction that causes the Python
interpreter to display a value on the screen.
\index{print statement}
\index{statement!print}

\item[operator:]  A special symbol that represents a simple computation like
addition, multiplication, or string concatenation.
\index{operator}

\item[value:]  One of the basic units of data, like a number or string,
that a program manipulates.
\index{value}

\item[type:] A category of values.  The types we have seen so far
are integers (type {\tt int}), floating-point numbers (type {\tt
float}), and strings (type {\tt str}).
\index{type}

\item[integer:] A type that represents whole numbers.
\index{integer}

\item[floating-point:] A type that represents numbers with fractional
parts.
\index{floating-point}

\item[string:] A type that represents sequences of characters.
\index{string}

\item[natural language:]  Any one of the languages that people speak that
evolved naturally.
\index{natural language}

\item[formal language:]  Any one of the languages that people have designed
for specific purposes, such as representing mathematical ideas or
computer programs; all programming languages are formal languages.
\index{formal language}

\item[token:]  One of the basic elements of the syntactic structure of
a program, analogous to a word in a natural language.
\index{token}

\item[syntax:] The rules that govern the structure of a program.
\index{syntax}

\item[parse:] To examine a program and analyze the syntactic structure.
\index{parse}

\item[bug:] An error in a program.
\index{bug}

\item[debugging:] The process of finding and correcting bugs.
\index{debugging}

\end{description}

\begin{description}

\item[问题求解 (problem solving)：]  将问题形式化、寻找并表达解决方案的过程。
\index{problem solving}  \index{问题求解}

\item[高级语言 (high-level language)：]  像Python这样被设计成人类容易阅读和编写的编程语言。
\index{high-level language}  \index{高级语言}

\item[低级语言 (low-level language)：]  被设计成计算机容易运行的编程语言；也被称为``机器语言 (machine language )'' 或``汇编语言'' (assembly language)。
\index{low-level language}  \index{低级语言}

\item[可移植性 (portability)：]  程序能够在多种计算机上运行的特性。
\index{portability}  \index{可移植性}

\item[解释器 (interpreter)：]  读取另一个程序并执行该程序的程序。
\index{interpret}  \index{解释器}

\item[提示符 (prompt)：] 解释器所显示的字符，表明已准备好接受用户的输入。
\index{prompt}  \index{提示符}

\item[程序 (program)：] 一组定义了计算内容的指令。
\index{program}  \index{程序}

\item[打印语句 (print statement)：]  使Python解释器在屏幕上显示某个值的指令。
\index{print statement}  \index{statement!print}
\index{打印语句}  \index{语句!打印}

\item[运算符 (operator)：]  代表类似加法、乘法或者字符串连接 (string concatenation) 等简单计算的特殊符号。
\index{operator}  \index{运算符}

\item[值 (value)：]  程序所处理数据的基本元素之一，例如数字或字符串。
\index{value}  \index{值}

\item[类型 (type)：] 值的类别。我们目前接触的类型有整型数（类型为 \lstinline{int}）、浮点数（类型为 \lstinline{float} ）和字符串（类型为 \lstinline{str} ）
\index{type}  \index{类型}

\item[整型数 (integer)：] 代表整数的类型。
\index{integer}  \index{整形数}

\item[浮点数 (floating-point)：] 代表一个有小数点的数字的类型。
\index{floating-point}  \index{浮点数}

\item[字符串 (string)：] A type that represents sequences of characters.
\index{string}  \index{字符串}

\item[自然语言 (natural language)：]  任何的人们日常使用的、由自然演变而来的语言。
\index{natural language}  \index{自然语言}

\item[形式语言 (formal language)：]  任何由人类为了某种目的而设计的语言，例如用来表示数学概念或者电脑程序；所有的编程语言都是形式语言。
\index{formal language}  \index{形式语言}

\item[记号 (token)：]  程序语法结构中的基本元素之一，与自然语言中的单词类似。
\index{token}  \index{记号}

\item[语法 (syntax)：] 规定了程序结构的规则。
\index{syntax}  \index{语法}

\item[解析 (parse)：] 阅读程序，并分析其语法结构的过程

\item[故障 (bug)：] 程序中的错误。
\index{bug}

\item[调试 (debugging)：] 寻找并解决错误的过程。
\index{debugging}  \index{调试}

\end{description}

\section{Exercises  |  练习}

\begin{exercise}

It is a good idea to read this book in front of a computer so you can
try out the examples as you go.

Whenever you are experimenting with a new feature, you should try
to make mistakes.  For example, in the ``Hello, world!'' program,
what happens if you leave out one of the quotation marks?  What
if you leave out both?  What if you spell {\tt print} wrong?
\index{error message}

This kind of experiment helps you remember what you read; it also
helps when you are programming, because you get to know what the error
messages mean.  It is better to make mistakes now and on purpose than
later and accidentally.

\begin{enumerate}

\item In a print statement, what happens if you leave out one
of the parentheses, or both?

\item If you are trying to print a string, what happens if you
leave out one of the quotation marks, or both?

\item You can use a minus sign to make a negative number like
{\tt -2}.  What happens if you put a plus sign before a number?
What about {\tt 2++2}?

\item In math notation, leading zeros are ok, as in {\tt 02}.
What happens if you try this in Python?

\item What happens if you have two values with no operator
between them?

\end{enumerate}


你最好在电脑前阅读此书，因为你可以随时测试书中的示例。

每当你试验一个新特性的时候，你应该试着去犯错。举个例子，在 “Hello, World!”程序中，如果你漏掉一个引号会发生什么情况？如果你去掉两个引号呢？如果你把 \lstinline{print} 写错了呢？
\index{error message}

这类试验能帮助你记忆读过的内容；对你平时编程也有帮助，因为你可以了解不同的错误信息代表的意思。现在故意犯错误，总胜过以后不小心犯错。

\begin{enumerate}

\item 在打印语句中，如果你去掉一个或两个括号，会发生什么？

\item 你想打印一个字符串，如果你去掉一个或两个引号，会发生什么？

\item 你可以使用减号创建一个负数，如 \lstinline{-2} 。如果你在一个数字前再加上个加号，会发生什么？ \lstinline{2++2} 会得出什么结果？

\item 在数学标记中，前导零 (leading zeros)没有问题，如 \lstinline{02} 。如果我们在 Python 中这样做，会发生什么？

\item 如果两个值之间没有运算符，又会发生什么？

\end{enumerate}

\end{exercise}



\begin{exercise}

Start the Python interpreter and use it as a calculator.

\begin{enumerate}

\item How many seconds are there in 42 minutes 42 seconds?

\item How many miles are there in 10 kilometers?  Hint: there are 1.61
  kilometers in a mile.

\item If you run a 10 kilometer race in 42 minutes 42 seconds, what is
  your average pace (time per mile in minutes and seconds)?  What is
  your average speed in miles per hour?

\index{calculator}
\index{running pace}

\end{enumerate}

启动Python解释器，把它当计算器使用。

\begin{enumerate}

\item 42分42秒一共有多少秒？

\item 10公里可以换算成多少英里？提示：一英里等于1.61公里。

\item 如果你花42分42秒跑完了10公里，你的平均配速（pace）\footnote{译注：配速（pace）是在马拉松运动的训练中常使用的一个概念，配速是速度的一种，是每公里所需要的时间。配速=时间/距离。}是多少（每英里耗时，分别精确到分和秒）？你每小时平均跑了多少英里（英里/时）？

\index{calculator}  \index{计算器}
\index{running pace}  \index{配速}

\end{enumerate}

\end{exercise}